Thesis: draft

### 2021-03-31 draft.tex
implemented lay-out feedback from Yolan, part 3

Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
\begin{equation}
	E \subseteq \{\{x, y\}| x, y \in V \}
\end{equation}
If self-loops are not allowed, we need the following condition on the elements of $E$
\begin{equation}
	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \}
\end{equation}
A graph with no self-loops is a simple graph. The nodes $x$ and $y$ of an edge $\{x, y\}$ are called the endpoints of the edge. It is possible that nodes are not joined by any edge, such nodes are called isolated or disconnected. If two or more edges have the same pair of endpoints, the graph is called a multi-graph. This thesis is not concerned with multi-graphs, nor are self-loops allowed. Graphs in which the edges have an orientation are called directed graphs. If the edges have weights $w_{ij}$, one speaks of a weighted graph.\\
A graph can be represented in different ways. One of the most common ways is by use of the adjacency matrix.\\

@@

The adjacency matrix \textbf{A} is a $N \times N$ matrix (for a graph with $N$ nodes) where the elements indicate whether two nodes are connected by an edge or not \cite{Mata2020}
\begin{equation}
	A_{ij} = 
	\begin{cases}
		1 & \text{if nodes $i$ and $j$ are connected}\\
		0 & \text{otherwise}
	\end{cases}
\end{equation}
For a simple, undirected graph, the adjacency matrix is symmetric ($A_{ij} = A_{ji}$) with zeros on the diagonal ($A_{ii} = 0$). Directed graphs can have asymmetric adjacency matrices. If we are dealing with weighted graphs, the elements of the adjacency matrix are the weights $w_{ij}$ of the edges
\begin{equation}
	A_{ij} = 
	\begin{cases}
		w_{ij} & \text{if nodes $i$ and $j$ are connected}\\
		0 & \text{otherwise}
	\end{cases}
\end{equation}
with, generally, $0 \leqslant w_{ij} \leqslant 1$ \cite{Mata2020}. \\
\subsection{Degree and degree distribution}

The adjacency matrix contains a lot of information such as the degree of a node. The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of node $i$. The degree can be easily obtained from the adjacency matrix \cite{Mata2020}
\begin{equation}
	k_i = \sum_{j=1}^N A_{ij}
\end{equation}
In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$. These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
\begin{equation}
	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}
\end{equation}
The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$. For weighted graphs the degree is easily generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
\begin{equation}
	s_i = \sum_{j=1}^N w_{ij}
\end{equation}
From the degree of the nodes in the network one can construct the degree distribution. The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$. The average degree can be obtained from the degree distribution \cite{Mata2020}
\begin{equation}
	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k)
\end{equation}
where $N$ is the total number of nodes in the network. The degree distribution also allows us to classify networks. The two most important classes are homogeneous and heterogeneous networks. Homogeneous networks have a bell curved degree distribution, eg. a Poisson distribution. In this case, most of the nodes have a degree close to the average degree $\left<k\right>$. Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$. % explain differences  more, consequences of power-law + need to include figure?
They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}. % correct statement? + talk about fact that av degree can diverge; explain scale free concept a bit more detailed
%maybe explain why we care about degree, centrality measure etc

@@

Another concept that is closely related to the degree of a node is the notion of assortative and disassortative networks. More precisely, the notion of assortativity and disassortativity has to do with degree correlations. \\
The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}. The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$ \cite{F.Costa2007}
\begin{equation}
	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)}
\end{equation}
The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
\begin{equation}
	k_{nn}(k) = \sum_{k'} k'P(k'|k)
\end{equation}
If there are no degree correlations, $k_{nn}$ is independent of $k$. Such a network is called a neutral network \cite{F.Costa2007}. If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative. If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
However, the terms assortativity and disassortativity are also often used without taking degree correlations into account. Instead, one may look at a particular node property to classify networks as either assortative or disassortative \cite{Thed2014}. In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.

@@

% explain length, diameter, average length,...
Several very useful measurements quantify the notion of length and distance in a network/graph. We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}. The length of a path is then defined as the number of edges in the path \cite{Goddard2010}. A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (note that the geodesic path is not necessarily unique) \cite{F.Costa2007}. The length of the geodesic path between the two nodes $i$ and $j$ is called the geodesic distance or shortest distance $d_{ij}$ \cite{F.Costa2007}. Geodesic distance is often just called distance \cite{Goddard2010}. The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\  
\newline
\newline
\textbf{Average distance}\\
\newline
The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
\begin{equation}\label{avdist}
	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
\end{equation}
This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}. This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes. These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}. Instead, another definition of average distance is used: the global efficiency \cite{F.Costa2007}
\begin{equation}
	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}
\end{equation}
This measurement quantifies the efficiency of the network in sending information between nodes \cite{F.Costa2007}. This assumes that the efficiency of sending information between two nodes is reciprocal to their distance. The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency \cite{F.Costa2007}
\begin{equation}
 	h = \frac{1}{E}
\end{equation}
\newline
\textbf{Diameter and radius}\\
\newline
The diameter $diam(G)$ of a network $G$ is the maximum distance between any pair of nodes \cite{Goddard2010}. The eccentricity of a node is the maximum distance from that node to any other node. The radius $rad(G)$ is the minimum eccentricity among all nodes of $G$. Note that the diameter is the maximum eccentricity among all nodes in $G$ \cite{Goddard2010}.
% finished?
\newpage
\subsection{Clustering coefficient and transitivity}

@@

	C_i = \frac{2n_i}{k_i(k_i-1)}

@@

	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i

@@

	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}}

@@

	N_{\Delta} &= \sum_{k > j > i} A_{ij}A_{ik}A_{jk} \\
	N_3 &= \sum_{k > j > i} (A_{ij}A_{ik} + A_{ji}A_{jk} + A_{ki}A_{kj})

@@

	T_i = \frac{N_{\Delta}(i)}{N_3(i)}

@@

	N_{\Delta}(i) &= \sum_{k > j} A_{ij}A_{ik}A_{jk} \\
	N_3(i) &= \sum_{k > j} A_{ij}A_{ik}

@@

Network communities are another feature that (complex) networks may posses. A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}. More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}. In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}. The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}. The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it. Several community definitions are consistent with the density hypothesis. Let us consider three possible definitions: maximum cliques, strong communitoes and weak communities \cite{Albert2016}.\\
\newline
\textbf{Maximum cliques}\\
\newline
A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}. A community based on this definition would then be the largest clique in the network. This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\   
\newpage
\noindent
\textbf{Strong communities}\\
\newline
A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}. Let us denote a community as $C$ and let $i$ be a node inside the community $C$. If we define the internal degree $k_i^{\textrm{int}}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{\textrm{ext}}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
\begin{equation}
	k_i^{\textrm{int}}(C) > k_i^{\textrm{ext}}(C) \qquad \forall i \in C
\end{equation}
\newline
\textbf{Weak communities}\\
\newline
A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}. We thus have the following condition for a weak community $C$ \cite{Albert2016}
\begin{equation}
	\sum_{i \in C} k_i^{\textrm{int}}(C) > \sum_{i \in C}k_i^{\textrm{ext}}(C)
\end{equation}

@@

	Q_c = \frac{1}{2L}\sum_{(i,j) \in C_c} (A_{ij} - p_{ij})

@@

	p_{ij} = \frac{k_ik_j}{2L}

@@

	Q_c = \frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2

@@

	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg]

@@

	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2}

@@

	k \leqslant \sqrt{2L}

@@

	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k}

@@

&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k}

@@

	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!}

@@

	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}

@@

	N \approx \left<k\right>^{d_{\textrm{max}}}

@@

	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}}

@@

	\left<d\right> \approx \frac{\ln{N}}{\ln{\left<k\right>}}

@@

In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering and short average distances. Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}. The construction goes as follows \cite{Watts1998}
\begin{enumerate}
	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$. Rewiring is done in such a way that self-loops and edge duplication are avoided
\end{enumerate}
The probability $\beta$ allows us to tune the graph between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}. Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$. The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<C(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
\begin{align}\label{C0}
	\left<C(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 && \left<d(0)\right> \approx \frac{N}{2K}
\end{align}
The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}. The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\left<C(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
\begin{align}
	\left<C(1)\right> \approx \frac{K}{N} \ll 1 && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}}
\end{align}
Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random graph (which is small), even for small $\beta$ \cite{Newman2000}. The clustering coefficient is approximately given by   \cite{Barrat1999}
\begin{equation}
	\left<C\right> \approx \left<C(0)\right>(1-\beta)^3
\end{equation} 
where $\left<C(0)\right>$ is given by Equation (\ref{C0}). This can be seen in the following way
\begin{itemize}
	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$
	\item The probability that edges are rewired back to each other is negligible
\end{itemize}
For rather small values of $\beta$, the clustering coefficient remains high. The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.

@@

	P(k) \sim k^{-\gamma}

@@

In the classical model, there are $N$ nodes/agents, each with an opinion $x_i$ and each node $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance/influence of the neighbors) \cite{Krause2002}. The most simple model has fixed weights $w$ and the updating rule is \cite{Krause2002}
\begin{equation}\label{refMod}
	x(t+1) = W x(t)
\end{equation}
where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
Friedkin and Johnsen proposed a variation of this model \cite{Friedkin1990}\cite{Friedkin1999}. The model they proposed assumes that an agent $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}. This variation thus has the following updating rule (in matrix notation) \cite{Krause2002}
\begin{equation}
	x(t+1) = Gx(0) + (\mathbb{1}-G)Wx(t)
\end{equation}
where $G$ is the diagonal matrix which contains the $g_i$ and $\mathbb{1}$ is the identity matrix. If all the $g_i$ are equal to zero, we obtain the model of Eq. (\ref{refMod}) \cite{Krause2002}.\\
A possible time variant extension is \cite{Krause2002}
\begin{equation}
	x(t+1) = W(t)x(t)
\end{equation}
where now the weights are time dependent. This model can be used for situations in which so called "hardening of positions" occur \cite{Krause2002}. This means that, over time, the agents put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}. 
% talk about extension http://jasss.soc.surrey.ac.uk/5/3/2/2.pdf

@@

	x_i(t+1) &= x_i(t) + \mu [x_j(t) - x_i(t)]\\
	x_j(t+1) &= x_j(t) + \mu [x_i(t) - x_j(t)]

>>>>

Networks or graphs are conceptually very simple and flexible objects that are made of a set of nodes/vertices $V$ and a set of edges/links $E$, where the elements of $E$ determine connections between elements of $V$ \cite{Costa2018}
\begin{equation}
	E \subseteq \{\{x, y\}| x, y \in V \}\ .
\end{equation}
If self-loops are not allowed, we need the following condition on the elements of $E$
\begin{equation}
	E \subseteq \{\{x, y\}| x, y \in V \text{and\ } x \neq y \} \ .
\end{equation}
A graph with no self-loops is a simple graph. The nodes $x$ and $y$ of an edge $\{x, y\}$ are called the endpoints of the edge. It is possible that nodes are not joined by any edge, such nodes are called isolated or disconnected. If two or more edges have the same pair of endpoints, the graph is called a multi-graph. This thesis is not concerned with multi-graphs, nor are self-loops allowed. Graphs in which the edges have an orientation are called directed graphs. If the edges have weights $w_{ij}$, one speaks of a weighted graph.\\
A graph can be represented in different ways. One of the most common ways is by use of the adjacency matrix.\\

@@

The adjacency matrix \textbf{A} is a $N \times N$ matrix (for a graph with $N$ nodes) where the elements indicate whether two nodes are connected by an edge or not \cite{Mata2020}
\begin{equation}
	A_{ij} = 
	\begin{cases}
		1 & \text{if nodes $i$ and $j$ are connected},\\
		0 & \text{otherwise}.
	\end{cases}
\end{equation}
For a simple, undirected graph, the adjacency matrix is symmetric ($A_{ij} = A_{ji}$) with zeros on the diagonal ($A_{ii} = 0$). Directed graphs can have asymmetric adjacency matrices. If we are dealing with weighted graphs, the elements of the adjacency matrix are the weights $w_{ij}$ of the edges
\begin{equation}
	A_{ij} = 
	\begin{cases}
		w_{ij} & \text{if nodes $i$ and $j$ are connected},\\
		0 & \text{otherwise},
	\end{cases}
\end{equation}
with, generally, $0 \leqslant w_{ij} \leqslant 1$ \cite{Mata2020}. \\
\subsection{Degree and degree distribution}

The adjacency matrix contains a lot of information such as the degree of a node. The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of node $i$. The degree can be easily obtained from the adjacency matrix \cite{Mata2020}
\begin{equation}
	k_i = \sum_{j=1}^N A_{ij}\ .
\end{equation}
In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$. These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
\begin{align}
	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \ , && k_i^{\text{out}} = \sum_{j=1}^N A_{ij} \ .
\end{align}
The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$. For weighted graphs the degree is easily generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
\begin{equation}
	s_i = \sum_{j=1}^N w_{ij} \ .
\end{equation}
From the degree of the nodes in the network one can construct the degree distribution. The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$. The average degree can be obtained from the degree distribution \cite{Mata2020}
\begin{equation}
	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k) \ ,
\end{equation}
where $N$ is the total number of nodes in the network. The degree distribution also allows us to classify networks. The two most important classes are homogeneous and heterogeneous networks. Homogeneous networks have a bell curved degree distribution, eg. a Poisson distribution. In this case, most of the nodes have a degree close to the average degree $\left<k\right>$. Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$. % explain differences  more, consequences of power-law + need to include figure?
They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}. % correct statement? + talk about fact that av degree can diverge; explain scale free concept a bit more detailed
%maybe explain why we care about degree, centrality measure etc

@@

Another concept that is closely related to the degree of a node is the notion of assortative and disassortative networks. More precisely, the notion of assortativity and disassortativity has to do with degree correlations. \\
The joint degree distribution $P(k, k')$ is the probability that an arbitrary edge connects a node with degree $k$ to a node with degree $k'$ \cite{F.Costa2007}. The conditional probability $P(k|k')$, which is another way to express dependencies between node degrees, is the probability that an arbitrary neighbor of a node with degree $k$ has a degree $k'$ \cite{F.Costa2007}
\begin{equation}
	P(k|k') = \frac{\left<k\right>P(k, k')}{kP(k)} \ .
\end{equation}
The average degree of the nearest neighbors of nodes with degree $k$ can be computed as \cite{F.Costa2007}
\begin{equation}
	k_{nn}(k) = \sum_{k'} k'P(k'|k) \ .
\end{equation}
If there are no degree correlations, $k_{nn}$ is independent of $k$. Such a network is called a neutral network \cite{F.Costa2007}. If $k_{nn}(k)$ is an increasing function of $k$, then nodes with a high degree tend to connect to nodes with a high degree and the network is classified as assortative. If instead $k_{nn}(k)$ is a decreasing function of $k$, nodes with a high degree tend to connect to nodes with a low degree and the network is called disassortative \cite{F.Costa2007}.\\
However, the terms assortativity and disassortativity are also often used without taking degree correlations into account. Instead, one may look at a particular node property to classify networks as either assortative or disassortative \cite{Thed2014}. In this way, an assortative network is a network where nodes tend to connect to other nodes with similar properties as themselves, whereas in a disassortative network nodes tend to connect to nodes with different properties \cite{Thed2014}.

@@

% explain length, diameter, average length,...
Several very useful measurements quantify the notion of length and distance in a network/graph. We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}. The length of a path is then defined as the number of edges in the path \cite{Goddard2010}. A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (note that the geodesic path is not necessarily unique) \cite{F.Costa2007}. The length of the geodesic path between the two nodes $i$ and $j$ is called the geodesic distance or shortest distance $d_{ij}$ \cite{F.Costa2007}. Geodesic distance is often just called distance \cite{Goddard2010}. The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\  
\newline
\textbf{Average distance}\\
\newline
The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
\begin{equation}\label{avdist}
	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij} \ .
\end{equation}
This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}. This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes. These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}. Instead, another definition of average distance is used: the global efficiency \cite{F.Costa2007}
\begin{equation}
	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}} \ .
\end{equation}
This measurement quantifies the efficiency of the network in sending information between nodes \cite{F.Costa2007}. This assumes that the efficiency of sending information between two nodes is reciprocal to their distance. The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency \cite{F.Costa2007}
\begin{equation}
 	h = \frac{1}{E} \ .
\end{equation}
\newline
\textbf{Diameter and radius}\\
\newline
The diameter $diam(G)$ of a network $G$ is the maximum distance between any pair of nodes \cite{Goddard2010}. The eccentricity of a node is the maximum distance from that node to any other node. The radius $rad(G)$ is the minimum eccentricity among all nodes of $G$. Note that the diameter is the maximum eccentricity among all nodes in $G$ \cite{Goddard2010}.
% finished?
\newpage
\subsection{Clustering coefficient and transitivity}

@@

	C_i = \frac{2n_i}{k_i(k_i-1)} \ ,

@@

	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i \ ,

@@

	T = \frac{3 \times \text{number of triangles in the network}}{\text{number of connected triples of nodes in the network}} \ .

@@

	N_{\Delta} &= \sum_{k > j > i} A_{ij}A_{ik}A_{jk} \ , \\
	N_3 &= \sum_{k > j > i} (A_{ij}A_{ik} + A_{ji}A_{jk} + A_{ki}A_{kj}) \ ,

@@

	T_i = \frac{N_{\Delta}(i)}{N_3(i)} \ ,

@@

	N_{\Delta}(i) &= \sum_{k > j} A_{ij}A_{ik}A_{jk} \ , \\
	N_3(i) &= \sum_{k > j} A_{ij}A_{ik} \ .

@@

Network communities are another feature that (complex) networks may posses. A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}. More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}. In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}. The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}. The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it. Several community definitions are consistent with the density hypothesis. Let us consider three possible definitions: maximum cliques, strong communitoes and weak communities \cite{Albert2016}.\\
\newline
\textbf{Maximum cliques}\\
\newline
A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}. A community based on this definition would then be the largest clique in the network. This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\   
\newpage
\noindent
\textbf{Strong communities}\\
\newline
A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}. Let us denote a community as $C$ and let $i$ be a node inside the community $C$. If we define the internal degree $k_i^{\textrm{int}}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{\textrm{ext}}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
\begin{equation}
	k_i^{\textrm{int}}(C) > k_i^{\textrm{ext}}(C) \ , \qquad \forall i \in C \ .
\end{equation}
\newline
\textbf{Weak communities}\\
\newline
A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}. We thus have the following condition for a weak community $C$ \cite{Albert2016}
\begin{equation}
	\sum_{i \in C} k_i^{\textrm{int}}(C) > \sum_{i \in C}k_i^{\textrm{ext}}(C) \ .
\end{equation}

@@

	Q_c = \frac{1}{2L}\sum_{(i,j) \in C_c} (A_{ij} - p_{ij}) \ ,

@@

	p_{ij} = \frac{k_ik_j}{2L} \ ,

@@

	Q_c = \frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \ ,

@@

	Q = \sum_{c = 1}^{n_c}\bigg[\frac{L_c}{L} - \bigg(\frac{k_c}{2L}\bigg)^2 \bigg] \ .

@@

	\Delta Q_{AB} = \frac{l_{AB}}{L} - \frac{k_Ak_B}{2L^2} \ ,

@@

	k \leqslant \sqrt{2L} \ ,

@@

	p_k = \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .

@@

&= \sum_{k=0}^{N-1} k \binom{N-1}{k}p^k (1-p)^{N-1-k} \ .

@@

	p_k = \mathrm{e}^{-\left<k\right>} \frac{\left<k\right>^k}{k!} \ .

@@

	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1} \ .

@@

	N \approx \left<k\right>^{d_{\textrm{max}}} \ ,

@@

	d_{\textrm{max}} \approx \frac{\ln{N}}{\ln{\left<k\right>}} \ .

@@

	\left<d\right> \approx \frac{\ln{N}}{\ln{\left<k\right>}} \ .

@@

In 1998, Duncan Watts and Steve Strogatz constructed a network model that generates random networks with small world properties, such as high clustering and short average distances. Their model achieves this by interpolating between a regular ring lattice and a random network \cite{Watts1998}. The construction goes as follows \cite{Watts1998}
\begin{enumerate}
	\item Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side)
	\item For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$. Rewiring is done in such a way that self-loops and edge duplication are avoided
\end{enumerate}
The probability $\beta$ allows us to tune the graph between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}. Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$. The first inequality assures sparse networks, whereas the second inequality guarantees that a random graph will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<C(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
\begin{align}\label{C0}
	\left<C(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 \ , && \left<d(0)\right> \approx \frac{N}{2K} \ .
\end{align}
The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}. The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
The random network ($\beta = 1$) on the other hand has a low clustering coefficient $\left<C(1)\right>$ and a small average path length $\left<d(1)\right>$ (see Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
\begin{align}
	\left<C(1)\right> \approx \frac{K}{N} \ll 1 \ , && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}} \ .
\end{align}
Watts and Strogatz then showed by numerical simulation that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random graph (which is small), even for small $\beta$ \cite{Newman2000}. The clustering coefficient is approximately given by   \cite{Barrat1999}
\begin{equation}
	\left<C\right> \approx \left<C(0)\right>(1-\beta)^3 \ ,
\end{equation} 
where $\left<C(0)\right>$ is given by Equation (\ref{C0}). This can be seen in the following way
\begin{itemize}
	\item The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$
	\item The probability that edges are rewired back to each other is negligible
\end{itemize}
For rather small values of $\beta$, the clustering coefficient remains high. The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances.

@@

	P(k) \sim k^{-\gamma} \ ,

@@

In the classical model, there are $N$ nodes/agents, each with an opinion $x_i$ and each node $i$ gives a weight $w_{ij}$ to all its neighbors (with $w_{ij} \geqslant 0$; this weight may for example represent the importance/influence of the neighbors) \cite{Krause2002}. The most simple model has fixed weights $w$ and the updating rule is \cite{Krause2002}
\begin{equation}\label{refMod}
	x(t+1) = W x(t) \ ,
\end{equation}
where $W$ is the fixed stochastic matrix which contains the weights $w_{ij}$, $x(t)$ is the opinion profile at time $t$ (this is a $N$-dimensional vector containing the opinions $x_i$ of each node at time $t$) and $x(t+1)$ is the opinion profile at time $t+1$.\\
Friedkin and Johnsen proposed a variation of this model \cite{Friedkin1990}\cite{Friedkin1999}. The model they proposed assumes that an agent $i$ adheres to its initial opinion to a certain degree $g_i$ and is influenced by others by a susceptibility of $1-g_i$ \cite{Krause2002}. This variation thus has the following updating rule (in matrix notation) \cite{Krause2002}
\begin{equation}
	x(t+1) = Gx(0) + (\mathbb{1}-G)Wx(t) \ ,
\end{equation}
where $G$ is the diagonal matrix which contains the $g_i$ and $\mathbb{1}$ is the identity matrix. If all the $g_i$ are equal to zero, we obtain the model of Eq. (\ref{refMod}) \cite{Krause2002}.\\
A possible time variant extension is \cite{Krause2002}
\begin{equation}
	x(t+1) = W(t)x(t) \ ,
\end{equation}
where now the weights are time dependent. This model can be used for situations in which so called "hardening of positions" occur \cite{Krause2002}. This means that, over time, the agents put more and more weight on their own opinion and less weight on the opinion of others \cite{Krause2002}. 
% talk about extension http://jasss.soc.surrey.ac.uk/5/3/2/2.pdf

@@

	x_i(t+1) &= x_i(t) + \mu [x_j(t) - x_i(t)] \ ,\\
	x_j(t+1) &= x_j(t) + \mu [x_i(t) - x_j(t)] \ .
